---
title: "L2_325_notes"
format: html
---

# Data from the table
```{r}
cotton_percentage <- c(14.9, 15.1, 15, 15, 15, 20, 20, 20, 20, 20, 25, 25, 25, 25, 25, 30, 30, 30, 30)
tensile_strength <- c(    7,    7,  9,  8, 10, 19, 20, 21, 20, 22, 21, 21, 17, 19, 20, 8, 7, 8, 10)
```

# Create the scatter plot
```{r}
plot(cotton_percentage, tensile_strength,
     main = "Figure 1.5 Scatter plot of tensile strength and cotton percentages",
     xlab = "Cotton Percentage",
     ylab = "Tensile Strength",
     pch = 16,       # Solid circles
     col = "blue")   # Color of points
```
```{r}
# Load necessary library
library(tibble)
# Create the dataframe (visual easy to understand, creating a tibble by Row)
data <- tribble(
~Company_A, ~Company_B, # The ~ symbol is used to specify the column names
8.8, 11.5,
8.4, 9.6,
7.3, 10.3,
8.1, 10.4,
8.2, 9.8,
6.5, 9.5,
7.8, 11.5,
6.4, 11.0,
9.2, 10.4,
7.0, 10.1
)
print(data)
library(tidyr)
long_data <- pivot_longer(data, cols = everything(), names_to = "Company", values_to = "Flexibility")
str(long_data)
dotchart(long_data$Flexibility,
         main = "Dotchart of Steel rods flexibility for two different companies",
         xlab = "Flexibility",
         groups = as.factor(long_data$Company),
         pch = 19,
         col = as.factor(long_data$Company))

#OR transform first to categoric value
str(long_data)
long_data$Company <- as.factor(long_data$Company)
str(long_data)
dotchart(long_data$Flexibility,
         main = "Dotchart of Steel rods flexibility for two different companies",
         xlab = "Flexibility",
         groups = long_data$Company,
         pch = 19,
         col = long_data$Company)
```
#Measures of Variablility
What is variability?
Sample range 
Calculate the range of the data 
variance is measure every data point to mean and sqaure it and sum it up and divide by n-1

very important for the solution of the problem


```{r}
data_point <- c(7.07, 7.00, 7.10, 6.97, 7.00, 7.02, 7.01, 7.00, 7.00, 7.00)
X_bar <- mean(data_point)
n <- length(data_point)
range(data_point)
#Example variance calculation
variance <- sum((data_point - X_bar)^2)/(n-1)
sd <- sqrt(variance)
variance
sd
summary(data_point)
#sort the data_point from small to big

sort(data_point)

# sort the data point from big to small
sort(data_point, decreasing = TRUE)
#sd_calculation

```

# new section 
```{r}
# Load necessary library
library(tibble)

# Create the dataframe (visual easy to understand, creating a tibble by Row)
data <- tribble(
  ~Company_A, ~Company_B, ~Company_C,# The ~ symbol is used to specify the column names
  8.8, 11.5, 3.4,
  8.4, 9.6,  3.2,
  7.3, 10.3, 3.1,
  8.1, 10.4, 3.3,
  8.2, 9.8,  3.2,
  6.5, 9.5,  3.1,
  7.8, 11.5, 3.4,
  6.4, 11.0, 3.3,
  9.2, 10.4, 3.2,
  7.0, 10.1, 3.1,
)

# Display the dataframe
print(data)
#print company A
data$Company_A

sd(data$Company_A)
mean(data$Company_A)
median(data$Company_A)

# plot the sd mean and median
plot(data$Company_A, main = "Company A", xlab = "Company A", ylab = "Flexibility", pch = 16, col = "blue")


```
```{r}
# histogram of the mean and median
compare_mean_median_A <- c(mean(data$Company_A), median(data$Company_A))
compare_mean_median_B <- c(mean(data$Company_B), median(data$Company_B))
print(compare_mean_median_A)
print(compare_mean_median_B)


```


what is data frame, is very much like excel.
create a data frame 
```{r}
a <- c(1,2,3,4,5)
b<- c(6,7,8,9,10)

data <- data.frame(a,b)
data
```
About libaries 'tidyr'
```{r}
library(tidyr)
long_data <- pivot_longer(data, cols = everything(), names_to = "Company", values_to = "Flexibility")
str(long_data)
view(long_data)
```

# ggplot
```{r}

library(ggplot2)
ggplot(long_data, aes(y = Company, x= Flexibility, color = Company)) + geom_point(size = 3, stroke = 1.5, shape = 21, fill = "white") + theme_minimal()
```
Example #2
```{r}
new_data_point <- tribble(
~Control,~Treatment,
  5,-6,
  6,4,
  -4,9,
  13,3,
  1,3,
  4,11,
  22,39,
  -9,4,
  9,6,
  4,6
)


```


